\section{Goals met} \label{sec:goals-met}
After analysing the experimental data gathered in the previous chapter and retrospectively looking back at the original objectives of this project, we can only conclude that all main objectives have successfully been achieved with the terminus of the project.

Some advanced functionality we had planned at the beginning ended up not being implemented, especially improvements addressing mostly user friendliness and ease of use on the software front.
But, as explained in a previous chapter, a proof-of-concept system is not a final product.
All the essential features and desirable characteristics have been implemented, per the project requirements.

The experimental results confirm the validity of the system proposed in this document and provide a good indication that future works based on the presented concept are likely to produce positive results.
Eventually, more advanced designs could be able to port the concept to other scientific domains other than automation engineering.

After taking a step back and taking a perspective look towards the project as a whole, the difficulties that we overcame, the results we gathered and the goals we had established in the beginning, one conclusion I have definitely reached is that the initial concept ideas would have been to complex for the task at hand.
The results obtained in such cases would have been influenced by many uncontrollable factors and would probably not reflect as much the concept we set out to explore: the network cycle time influence in control applications using Real-Time Ethernet networks.

Although simple, the developed system and concept have been able to produce results with minimum external influence.
As said before, most times implementing and designing less can mean much more, and the results are proof of that.
A barebones Distributed Control System has allowed us to explore the influence of the network cycle time on control systems.

As explained in the introduction sections, we aim at providing a solid foundation on which more advanced concepts and systems ban be built upon.
We have followed and fulfilled all requirements that have been presented alongside the concept and we can only hope that such requirements have been correctly evaluated, because now only time will tell if the proposed system is solid, robust and flexible enough to be built upon.
